*Prochlorococcus marinus* responses to light and oxygen

### 2024-03-08 Code/Manuscript_O2.Rmd
added orcid numbersin readme

HL clades are more recently branching lineages, with reduced genome sizes in comparison to LL clades. High-Light clades are typically dominant picophytoplankters in near-surface, oligotrophic waters, characterized by high light levels and subdivided into clades based on iron adaptation [@johnsonNichePartitioningProchlorococcus2006; @zinserInfluenceLightTemperature2007; @kentParallelPhylogeographyProchlorococcus2019]. Clade HLI, represented by cultured strain MED4, is adapted to high iron, and low temperatures, and originated from 5 m depth in the Mediterranean Sea [@partenskyProchlorococcusMarinePhotosynthetic1999]. Clade HLII, adapted to high iron, and high temperatures, is the most abundant *P. marinus* clade in the North Atlantic and North Pacific Oceans, often constituting over 90% of the total population [@partenskyProchlorococcusMarinePhotosynthetic1999], and are most numerous around 50 m depth [@partenskyProchlorococcusMarinePhotosynthetic1999]. Clade HLIII/IV is adapted to low iron [@kentParallelPhylogeographyProchlorococcus2019; @johnsonNichePartitioningProchlorococcus2006; @zinserInfluenceLightTemperature2007].

@@

*Prochlorococcus* genomes encode 4 FtsH proteins, henceforth referred to as FtsH1-4, homologs to the characterized FtsH isoforms of the model freshwater cyanobacterium *Synechocystis* FtsH, and with presumably parallel functions (Table \@ref(tab:FtsHTable)). Upon a shift to higher light HLI MED4 upregulated expression of FtsH1 and FtsH2 [@bonisteelStrainSpecificDifferences2018], homologs to the *Synechocystis* *slr*0228 and *slr*1604, implicated in PSII repair. In contrast, representative LLIV strain MIT9313 showed less overall expression of the FtsH proteases, and thus has fewer FtsH serving each photosystem.  Furthermore, MIT9313 expressed primarily FtsH3, homologous to *Synechocystis* *slr*1463, possibly involved in PSI biogenesis, and FtsH expression did not increase in response to light stress in MIT9313 [@bonisteelStrainSpecificDifferences2018]. Through adaptation to steady low light, clade LLIV *Prochlorococcus* instead allocate resources to processes other than dynamic regulation of PSII repair.

@@

*Prochlorococcus marinus* MED4 requires more than 4 h of light per day; thus this strain will not exploit habitats typical of temperate winter or light attenuated depths, even if water temperatures warm into the clade HLI tolerance range. MED4 is also excluded from the lowest oxygen habitats of 2.5 µM O~2~, but can, grow under OMZ regions with 25 µM O~2~. Genomic (Figure \@ref(fig:ProchlorococcusEnzymeKms)) and transcriptional analyses [@bergResponsesPsbAHli2011]  suggest MED4 is excluded from growth below ~ 25 µM O~2~ because it relies upon a ubiquinol oxidase, non-electrogenic, to maintain oxidation/reduction balance in the intersystem electron transport chain, with a K~M~ for  [O~2~] of ~25 µM O~2~.  On the other hand, MED4 shows inducible expression of FtsH isoforms [@bonisteelStrainSpecificDifferences2018], to counter photoinactivation of PSII under higher PAR and [O~2~] environments. However, photoinactivation, does impose an increased cost of growth upon MED4, since growth under red light, to lower photoinactivation of PSII [@murphyPhotoinactivationPhotosystemII2017], allows MED4 to achieve faster growth per absorbed photon than growth under blue light. TARA Oceans Project data [@delmontLinkingPangenomesMetagenomes2018] indeed reported presence of *P. marinus* MED4-like genomes at depths ranging from 5 m to 90 m, representing high to low blue light levels, in the Pacific South East Ocean. Delmont and Eren [@delmontLinkingPangenomesMetagenomes2018] did not analyze data from depths beyond the subsurface chlorophyll maximum layer, nor did they report[O~2~] at depth. Our growth findings are  consistent with Figure \@ref(fig:CladeProchloroPhotosynthDepthO2) showing PSII proteins annotated as MED4, clade HLI, at depths up to 200 meters, with O~2~ of ~15 µM.

>>>>

High-Light clades are more recently branching lineages, with reduced genome sizes in comparison to LL clades. High-Light clades are typically dominant picophytoplankters in near-surface, oligotrophic waters, characterized by high light levels and subdivided into clades based on iron adaptation [@johnsonNichePartitioningProchlorococcus2006; @zinserInfluenceLightTemperature2007; @kentParallelPhylogeographyProchlorococcus2019]. Clade HLI, represented by cultured strain MED4, is adapted to high iron, and low temperatures, and originated from 5 m depth in the Mediterranean Sea [@partenskyProchlorococcusMarinePhotosynthetic1999]. Clade HLII, adapted to high iron, and high temperatures, is the most abundant *P. marinus* clade in the North Atlantic and North Pacific Oceans, often constituting over 90% of the total population [@partenskyProchlorococcusMarinePhotosynthetic1999], and are most numerous around 50 m depth [@partenskyProchlorococcusMarinePhotosynthetic1999]. Clade HLIII/IV is adapted to low iron [@kentParallelPhylogeographyProchlorococcus2019; @johnsonNichePartitioningProchlorococcus2006; @zinserInfluenceLightTemperature2007].

@@

*Prochlorococcus* genomes encode 4 FtsH proteins, henceforth referred to as FtsH1-4, homologs to the characterized FtsH isoforms of the model freshwater cyanobacterium *Synechocystis* FtsH, and with presumably parallel functions (Table \@ref(tab:FtsHTable)). Upon a shift to higher light, HLI MED4 upregulated expression of FtsH1 and FtsH2 [@bonisteelStrainSpecificDifferences2018], homologs to the *Synechocystis* *slr*0228 and *slr*1604, implicated in PSII repair. In contrast, representative LLIV strain MIT9313 showed less overall expression of the FtsH proteases, and thus has fewer FtsH serving each photosystem.  Furthermore, MIT9313 expressed primarily FtsH3, homologous to *Synechocystis* *slr*1463, possibly involved in PSI biogenesis, and FtsH expression did not increase in response to light stress in MIT9313 [@bonisteelStrainSpecificDifferences2018]. Through adaptation to steady low light, clade LLIV *Prochlorococcus* instead allocate resources to processes other than dynamic regulation of PSII repair.

@@

*Prochlorococcus marinus* MED4 requires more than 4 h of light per day; thus this strain will not exploit habitats typical of temperate winter or light attenuated depths, even if water temperatures warm into the clade HLI tolerance range. MED4 is also excluded from the lowest oxygen habitats of 2.5 µM O~2~, but can, grow under OMZ regions with 25 µM O~2~. Genomic (Figure \@ref(fig:ProchlorococcusEnzymeKms)) and transcriptional analyses [@bergResponsesPsbAHli2011]  suggest MED4 is excluded from growth below ~ 25 µM O~2~ because it relies upon a ubiquinol oxidase, non-electrogenic, to maintain oxidation/reduction balance in the intersystem electron transport chain, with a K~M~ for  [O~2~] of ~25 µM O~2~.  On the other hand, MED4 shows inducible expression of FtsH isoforms [@bonisteelStrainSpecificDifferences2018], to counter photoinactivation of PSII under higher PAR and [O~2~] environments. However, photoinactivation imposes an increased cost of growth upon MED4, since growth under red light, to lower photoinactivation of PSII [@murphyPhotoinactivationPhotosystemII2017], allows MED4 to achieve faster growth per absorbed photon than growth under blue light. TARA Oceans Project data [@delmontLinkingPangenomesMetagenomes2018] indeed reported presence of *P. marinus* MED4-like genomes at depths ranging from 5 m to 90 m, representing high to low blue light levels, in the Pacific South East Ocean. Delmont and Eren [@delmontLinkingPangenomesMetagenomes2018] did not analyze data from depths beyond the subsurface chlorophyll maximum layer, nor did they report[O~2~] at depth. Our growth findings are  consistent with Figure \@ref(fig:CladeProchloroPhotosynthDepthO2) showing PSII proteins annotated as MED4, clade HLI, at depths up to 200 meters, with O~2~ of ~15 µM.
